Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-04-11 Moltres.tex
Add conclusion section.

This work does not introduce new science relative to the current literature so
much as it introduces a new open tool, Moltres
\cite{lindsayad_arfc/moltres_nodate}, for \gls{MSR} simulation. Moltres is an
open source code licensed under LGPL2 terms that aims to be a tool that can be
freely used, studied, critiqued, and developed by the \gls{MSR}
community. Though Moltres is devoted to an accurate description of \gls{MSR}
physics, its openness is its defining characteristic and one that actually
promotes the former quality through transparency and easy peer review. In the
era of github and global development, open and modern software practices must be
employed in order for nuclear engineering simulation capability to keep pace
with that of fields like astronomy whose multitude of open simulation tools
(Enzo, Gadget, Flash, Gizmo, Ramses, Mesa) allows them to be the darlings of
scientific visualization and documentaries. In that vein Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another LGPL
code that itself leans on LibMesh \cite{kirk2006libmesh}, a LGPL finite element
library, and PetSc \cite{petsc-user-ref}, a BSD toolkit for solving nonlinear
equations stemming from discretization of PDEs. MOOSE and LibMesh handle
translation of application (Moltres) defined weak PDE forms into residual and
Jacobian functions that are the inputs into PetSc Newton-Raphson solution
routines. All codes use MPI for parallel communication and are easily deployed
on massively-parallel cluster-computing platforms. MOOSE applications by default
use monolithic and implicit methods that are ideal for closely-coupled and
multi-scale physics. These methods are ideal for the model problem described in
this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it more
extensible to future problems.

@@

The model is implemented using the MOOSE \cite{gaston_physics-based_2015} application code
Moltres. \cite{lindsayad_arfc/moltres_nodate} Building on the massively
parallelizable MOOSE framework allows Moltres to be run on super-computing
platforms like \gls{NCSA}'s Blue Waters. For some three-dimensional simulations
(results not discussed here), the number of elements in the mesh and total
number of degrees of freedom exceed one million and ten million respectively. To
handle problems of this size, Moltres has been run on up to 608 cores. However,
reducing the problem dimension from three to two and using a structured mesh
that can be much more coarse in the axial direction allows problem solution to
be carried out on a single core in under five minutes.

The Moltres code-base currently includes kernels and boundary conditions for
solving for neutron fluxes, temperature, and precursor concentrations. In MOOSE
jargon, kernels are C++ classes that contain methods for computing residual and
Jacobian contributions corresponding to individual piecies of governing
equations. Developing the code-base in this way allows very modular construction
of equation systems; e.g. the same kernel that is used to represent heat
conduction can also be used to represent generic chemical diffusion. Moltres
also features neutron and precursor ``actions.'' These actions enable automatic
construction of the systems of equations for an arbitrary number of neutron and
precursor groups. This means that as long as group constants are provided in an appropriate
tabular form, a user only has to modify a couple of lines in a Moltres input
file to change from say two to fourty-four neutron groups.

@@

\begin{table}[htpb]
  \begin{center}
    \begin{tabular}{l | r}
      Component & Mass Fraction\\\hline\hline
      Li-7 & .1090\\
      Li-6 & 5$\cdot$10$^{-6}$\\
      F-19 & .6680\\
      Be-9 & .0627\\
      U-235 & .0167\\
      U-238 & .0344\\
    \end{tabular}
  \end{center}
  \caption{Fuel salt composition based on enriched Uranium \gls{MSRE} design \cite{robertson_msre_1965}
  \label{table:comp}
\end{table}

@@

The effect of the advecting fuel is seen in {fig:temp} where the temperature
rises steadily with increasing z-coordinate. As one might expect, the temperature
gradient is negative in the radial direction.

@@

>>>>

This work does not introduce new science relative to the current literature so
much as it introduces a new open tool, Moltres
\cite{lindsay_arfc/moltres_nodate}, for \gls{MSR} simulation. Moltres is an
open source code licensed under \gls{LGPL} terms that aims to be a tool that can be
freely used, studied, critiqued, and developed by the \gls{MSR}
community. Though Moltres is devoted to an accurate description of \gls{MSR}
physics, its openness is its defining characteristic and one that actually
promotes the former quality through transparency and easy peer review. In the
era of github and global development, open and modern software practices must be
employed in order for nuclear engineering simulation capability to keep pace
with that of fields like astronomy whose multitude of open simulation tools
(Enzo, Gadget, Flash, Gizmo, Ramses, Mesa) allows them to be the darlings of
scientific visualization and documentaries. In that vein Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another \gls{LGPL}
code that itself leans on LibMesh \cite{kirk2006libmesh}, a \gls{LGPL} finite element
library, and PetSc \cite{petsc-user-ref}, a \gls{BSD}-licensed toolkit for solving nonlinear
equations stemming from discretization of PDEs. \gls{MOOSE} and LibMesh handle
translation of application (Moltres) defined weak PDE forms into residual and
Jacobian functions that are the inputs into PetSc Newton-Raphson solution
routines. All codes use MPI for parallel communication and are easily deployed
on massively-parallel cluster-computing platforms. \gls{MOOSE} applications by default
use monolithic and implicit methods that are ideal for closely-coupled and
multi-scale physics. These methods are ideal for the model problem described in
this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it more
extensible to future problems.

@@

The model is implemented using the \gls{MOOSE} \cite{gaston_physics-based_2015} application code
Moltres. \cite{lindsay_arfc/moltres_nodate} Building on the massively
parallelizable \gls{MOOSE} framework allows Moltres to be run on super-computing
platforms like \gls{NCSA}'s Blue Waters. For some three-dimensional simulations
(results not discussed here), the number of elements in the mesh and total
number of degrees of freedom exceed one million and ten million respectively. To
handle problems of this size, Moltres has been run on up to 608 cores. However,
reducing the problem dimension from three to two and using a structured mesh
that can be much more coarse in the axial direction allows problem solution to
be carried out on a single core in under five minutes.

The Moltres code-base currently includes kernels and boundary conditions for
solving for neutron fluxes, temperature, and precursor concentrations. In \gls{MOOSE}
jargon, kernels are C++ classes that contain methods for computing residual and
Jacobian contributions corresponding to individual piecies of governing
equations. Developing the code-base in this way allows very modular construction
of equation systems; e.g. the same kernel that is used to represent heat
conduction can also be used to represent generic chemical diffusion. Moltres
also features neutron and precursor ``actions.'' These actions enable automatic
construction of the systems of equations for an arbitrary number of neutron and
precursor groups. This means that as long as group constants are provided in an appropriate
tabular form, a user only has to modify a couple of lines in a Moltres input
file to change from say two to fourty-four neutron groups.

@@

\begin{table}[htpb]
  \begin{center}
    \begin{tabular}{l | r}
      Component & Mass Fraction\\\hline\hline
      Li-7 & .1090\\
      Li-6 & 5$\cdot$10$^{-6}$\\
      F-19 & .6680\\
      Be-9 & .0627\\
      U-235 & .0167\\
      U-238 & .0344\\
    \end{tabular}
  \end{center}
  \caption{Fuel salt composition based on enriched Uranium \gls{MSRE} design \cite{robertson_msre_1965}}
  \label{table:comp}
\end{table}

@@

The effect of the advecting fuel is seen in \cref{fig:temp} where the temperature
rises steadily with increasing z-coordinate. As one might expect, the temperature
gradient is negative in the radial direction.

@@

\section{Conclusion}

This work introduces the open source \gls{MSR} simulation code Moltres. Moltres
solves arbitrary-group neutron diffusion, temperature, and precursor governing
equations in anywhere from one to three dimensions and can be deployed on an
arbitrary number of processing units. The model problem presented here has a
2D-axisymmetric geometry with heterogeneous group constants for fuel and
moderator regions. Fuel volume fraction and fuel salt composition are based on
the \gls{MSRE}. Neutron fluxes show expected cosinusoidal shapes in radial and
axial directions with visible striations between fuel and moderator regions. The
fast group flux is enhanced in fuel regions while the thermal group flux in
enhanced in moderator regions. Due to advection the temperature profile
increases monotonically in the direction of salt flow. The role of advection is
also seen in precursor concentrations. Long lived precursors exhibit maximum
concentrations at the core outlet. As the decay constant increases across
precursor groups the maximum concentrations moves towards the reactor center
where the precursor production rate is maximum. Future Moltres work includes
generating a high-fidelity 3D model as well as investigating various transient
accident scenarios.

### 2017-04-11 acros.tex
Add conclusion section.

@@

\newacronym{CEA}{CEA}{Commissariat \`a l'\'Energie Atomique et aux \'Energies Alternatives}

@@

@@

@@

>>>>

\newacronym{BSD}{BSD}{Berkeley Software Distribution}

@@

\newacronym{CEA}{CEA}{Commissariat \`a l'\'Energie Atomique et aux \'Energies
Alternatives}
\newacronym{CG}{CG}{Continuous Galerkin}

@@

\newacronym{DG}{DG}{Discontinuous Galerkin}

@@

\newacronym{LGPL}{LGPL}{Lesser GNU Public License}

@@

\newacronym{PetSc}{PetSc}{Portable, Extensible Toolkit for Scientific Computation}
